\textbf{SPECFEM3D}\newline \textbf{Cartesian}\newline \textbf{User Manual}

### 2022-11-12 doc/USER_MANUAL/01_introduction.tex
updates tex files (\\ to \newline) for better pandoc conversions to wiki pages

The software package SPECFEM3D Cartesian simulates seismic wave propagation
at the local or regional scale and performs full waveform imaging (FWI) or adjoint tomography based upon the spectral-element method
(SEM). The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
which can easily be made discontinuous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
it is then close to a particular case of the discontinuous Galerkin
technique \citep{ReHi73,LeRa74,Arn82,JoPi86,BoMaHe91,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10},
with optimized efficiency because of its tensorized basis functions
\citep{WiStBuGh10,AcKo11}. In particular, it can accurately handle
very distorted mesh elements \citep{OlSe11}.\\

In fluids, when gravity is turned off, SPECFEM3D uses the classical
linearized Euler equation; thus if you have sharp local variations of
density in the fluid (highly heterogeneous fluids
in terms of density) or if density becomes extremely small in some
regions of your model (e.g. for upper-atmosphere studies), before using
the code please make sure the linearized Euler equation is a valid
approximation in the case you want to study, and/or see if you should turn gravity on.
For more details on that see e.g. \cite{COA2011}.\\

It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08,AiWa09,AiWa10,MeStTh12}.
The spectral element approach admits spectral rates of convergence
and allows exploiting $hp$-convergence schemes. It is also very well
suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10}
as well as on clusters of GPU accelerating graphics cards \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
Tensor products inside each element can be optimized to reach very
high efficiency \citep{DeFiMu02}, and mesh point and element numbering
can be optimized to reduce processor cache misses and improve cache
reuse \citep{KoLaMi08a}. The SEM can also handle triangular (in 2D)
or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06}
as well as mixed meshes, although with increased cost and reduced
accuracy in these elements, as in the discontinuous Galerkin method.\\

Note that in many geological models in the context of seismic wave
propagation studies (except for instance for fault dynamic rupture
studies, in which very high frequencies or supershear rupture need
to be modeled near the fault, see e.g. \citet{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
a continuous formulation is sufficient because material property contrasts
are not drastic and thus conforming mesh doubling bricks can efficiently
handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\\

For a detailed introduction to the SEM as applied to regional seismic
wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07}
and in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
A detailed theoretical analysis of the dispersion
and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07}, \citet{SeOl08} and \citet{MeStTh12}.\\

Effects due to lateral variations in compressional-wave speed, shear-wave
speed, density, a 3D crustal model, topography and bathymetry are
included. The package can accommodate full 21-parameter anisotropy
(see~\citet{ChTr07}) as well as lateral variations in attenuation
\citep{SaKoTr10}. Adjoint capabilities and finite-frequency kernel
simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\\

The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
and has been successfully adapted to address problems in seismic wave
propagation. Early seismic wave propagation applications of the SEM,
utilizing Legendre basis functions and a perfectly diagonal mass matrix,
include \citet{CoJoTo93}, \citet{Kom97}, \citet{FaMaPaQu97}, \citet{CaGa97},
\citet{KoVi98} and \citet{KoTr99}, whereas applications involving
Chebyshev basis functions and a non-diagonal mass matrix include \citet{SePr94},
\citet{PrCaSe94} and \citet{SePrPr95}.
In the Legendre version that we use in SPECFEM the mass matrix is purposely slightly inexact but diagonal (but can be made exact if needed, see \cite{Teu15}),
while in the Chebyshev version it is exact but non diagonal.\\

Beware that, in a spectral-element method, some spurious modes (that have some similarities with classical so-called "Hourglass modes" in finite-element techniques,
although in the SEM they are not zero-energy modes) can appear in some (but not all) cases in the spectral element in which the source is located.
Fortunately, they do not propagate away from the source element.
However, this means that if you put a receiver in the same spectral element as a source, the recorded signals may in some cases be wrong, typically exhibiting some spurious
oscillations, which are often even non causal.
If that is the case, an easy option is to slightly change the mesh in the source region in order to get rid of these Hourglass-like spurious modes,
as explained in \cite{DuLiScGa14}, in which this phenomenon is described in details, and in which practical solutions to avoid it are suggested.\\

All SPECFEM3D software is written in Fortran2003 with full portability
in mind, and conforms strictly to the Fortran2003 standard. It uses
no obsolete or obsolescent features of Fortran. The package uses parallel
programming based upon the Message Passing Interface (MPI) \citep{GrLuSk94,Pac97}.\\

SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
conference in Phoenix, Arizona (USA) (see \citet{KoTsChTr03} and
\url{www.sc-conference.org/sc2003/nrfinalaward.html}). It was a finalist
again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors
of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories
(USA) \citep{CaKoLaTiMiLeSnTr08}. It also won the BULL Joseph Fourier
supercomputing award in 2010.\\

It reached the sustained one petaflop performance level for the first time in February 2013
on the Blue Waters Cray supercomputer at the National Center for Supercomputing Applications (NCSA), located at the University of Illinois at Urbana-Champaign (USA).\\

This new release of the code includes Convolution or
Auxiliary Differential Equation Perfectly Matched absorbing Layers
(C-PML or ADE-PML) \citep{MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10,KoMa07}.
It also includes support for GPU graphics card acceleration \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.\\

The next release of the code will use the PT-SCOTCH parallel and threaded
version of SCOTCH for mesh partitioning instead of the serial version.\\

@@

\subsection*{Announcements}
\begin{itemize}
\item {\bf FWI}: SPECFEM3D can now perform full waveform inversion (FWI), i.e. invert for models in an iterative fashion, and it can also perform source inversions in a constant model;
please refer to the two new directories, \texttt{inverse\_problem\_for\_model} and \texttt{inverse\_problem\_for\_source}, and the README files they contain. For FWI inversions for the model, also refer to the new examples provided in
directory \texttt{EXAMPLES}.\\

\item {\bf External coupling}: SPECFEM3D can now perform coupling with an external code (DSM, AxiSEM or FK) based on a database of displacement vectors and traction vectors
on the outer edges of the mesh created once and for all (see \cite{MoChKoFu13,MoChKoWa15,WaChMoKoMo16,ToChKoBaLi14,ToKoTsHuChBaLi14}, and if you use that feature of the code please cite at least one of these papers). \\

To use coupling with FK, just use the set of parameters that is in the \texttt{DATA/Par\_file} input file of the code: \\
%
\begin{verbatim}
#-----------------------------------------------------------
#
# Coupling with an injection technique (DSM, AxiSEM, or FK)
#
#-----------------------------------------------------------
COUPLE_WITH_INJECTION_TECHNIQUE = .false.
INJECTION_TECHNIQUE_TYPE        = 3       # 1 = DSM, 2 = AxiSEM, 3 = FK
MESH_A_CHUNK_OF_THE_EARTH       = .false.
TRACTION_PATH                   = ./DATA/AxiSEM_tractions/3/
FKMODEL_FILE                    = FKmodel
RECIPROCITY_AND_KH_INTEGRAL     = .false. # does not work yet
\end{verbatim}
That part (coupling with FK) is actively maintained and works fine. See e.g. \href{https://doi.org/10.1002/2014GL061644}{GRL\_Ping\_Tong\_2014} for some examples.
There is also an example that is provided with the code:\\
\texttt{specfem3d/EXAMPLES/small\_example\_coupling\_FK\_specfem}. \\

Regarding coupling with DSM, that part is not actively maintained any more, but it is still included in the code, you may have to test it again and make minor adjustments if needed. The necessary tools are in directory \texttt{specfem3d/EXTERNAL\_PACKAGES\_coupled\_with\_SPECFEM3D/DSM\_for\_SPECFEM3D}, and there is a \texttt{README} file in \texttt{specfem3d/EXTERNAL\_PACKAGES\_coupled\_with\_SPECFEM3D} that should be more or less up-to-date (there are about four steps to follow in total, the first one being creating the database of DSM tractions and displacements on the edges of the coupling box).
See e.g. \href{https://doi.org/10.1093/gji/ggs006}{GJI\_Vadim\_2013} and \href{https://doi.org/10.1093/gji/ggv189}{GJI\_Vadim\_2015} for some examples.\\

\item {\bf Gravity}: SPECFEM3D can now perform gravity field calculations in addition (or instead of) seismic wave propagation only.
See flag \texttt{GRAVITY\_INTEGRALS} in file \texttt{setup/constants.h}. Please also refer to \href{https://doi.org/10.1093/gji/ggx010}{GJI\_Martin\_gravimetry\_2017}.
And yes, that is the reason why Dimitri added a falling apple on the cover of the manual :-).
Note that SPECFEM3D can also model transient gravity perturbations induced by earthquake rupture, as developed and explained in \cite{HaAmBaChMoSoWh2015}.
These are two different things, and both are implemented and avaible in SPECFEM3D.
To use the second feature, please refer to \texttt{doc/how\_to/README\_gravity\_perturbation.txt}.\\

@@

You can find all the references below in \BibTeX format in file \texttt{doc/USER\_MANUAL/bibliography.bib}.\\

If you use SPECFEM3D Cartesian for your own research, please cite
at least one of the following articles:
\begin{description}
\item [{Numerical simulations in general}] ~\\
 Forward and adjoint simulations are described in detail in \citet{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99}
or \citet{KoVi98}. Additional aspects of adjoint simulations are
described in \citet{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
Domain decomposition is explained in detail in \citet{MaKoBlLe08},
and excellent scaling up to 150,000 processor cores is shown for instance
in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11},
\item [{GPU computing}] ~\\
 Computing on GPU graphics cards for acoustic or seismic wave propagation
applications is described in detail in \citet{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
\end{description}
\noindent If you use this new version, which has non blocking
MPI for much better performance for medium or large runs, please cite
at least one of these six articles, in which results of non blocking
MPI runs are presented: \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\\

@@

\noindent If you work on geophysical applications, you may be interested
in citing some of these application articles as well, among others:
\begin{description}
\item [{Southern California simulations}] ~\\
 \citet{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.

@@

\item [{Anisotropy}] ~\\
 \citet{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
\item [{Attenuation}] ~\\
 \citet{SaKoTr10,KoTr02a,KoTr99}.
\item [{Topography}] ~\\
 \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
\end{description}
The corresponding Bib\TeX{} entries may be found in file \texttt{doc/USER\_MANUAL/bibliography.bib}.

>>>>

The software package SPECFEM3D Cartesian simulates seismic wave propagation
at the local or regional scale and performs full waveform imaging (FWI) or adjoint tomography based upon the spectral-element method
(SEM). The SEM is a continuous Galerkin technique \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11},
which can easily be made discontinuous \citep{BeMaPa94,Ch00,KoWoHu02,ChCaVi03,LaWaBe05,Kop06,WiStBuGh10,AcKo11};
it is then close to a particular case of the discontinuous Galerkin
technique \citep{ReHi73,LeRa74,Arn82,JoPi86,BoMaHe91,FaRi99,HuHuRa99,CoKaSh00,GiHeWa02,RiWh03,MoRi05,GrScSc06,AiMoMu06,BeLaPi06,DuKa06,DeSeWh08,PuAmKa09,WiStBuGh10,DeSe10,EtChViGl10},
with optimized efficiency because of its tensorized basis functions
\citep{WiStBuGh10,AcKo11}. In particular, it can accurately handle
very distorted mesh elements \citep{OlSe11}.\newline

In fluids, when gravity is turned off, SPECFEM3D uses the classical
linearized Euler equation; thus if you have sharp local variations of
density in the fluid (highly heterogeneous fluids
in terms of density) or if density becomes extremely small in some
regions of your model (e.g. for upper-atmosphere studies), before using
the code please make sure the linearized Euler equation is a valid
approximation in the case you want to study, and/or see if you should turn gravity on.
For more details on that see e.g. \cite{COA2011}.\newline

It has very good accuracy and convergence properties \citep{MaPa89,SePr94,DeFiMu02,Coh02,DeSe07,SeOl08,AiWa09,AiWa10,MeStTh12}.
The spectral element approach admits spectral rates of convergence
and allows exploiting $hp$-convergence schemes. It is also very well
suited to parallel implementation on very large supercomputers \citep{KoTsChTr03,TsKoChTr03,KoLaMi08a,CaKoLaTiMiLeSnTr08,KoViCh10}
as well as on clusters of GPU accelerating graphics cards \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
Tensor products inside each element can be optimized to reach very
high efficiency \citep{DeFiMu02}, and mesh point and element numbering
can be optimized to reduce processor cache misses and improve cache
reuse \citep{KoLaMi08a}. The SEM can also handle triangular (in 2D)
or tetrahedral (in 3D) elements \citep{WinBoyd96,TaWi00,KoMaTrTaWi01,Coh02,MeViSa06}
as well as mixed meshes, although with increased cost and reduced
accuracy in these elements, as in the discontinuous Galerkin method.\newline

Note that in many geological models in the context of seismic wave
propagation studies (except for instance for fault dynamic rupture
studies, in which very high frequencies or supershear rupture need
to be modeled near the fault, see e.g. \citet{BeGlCrViPi07,BeGlCrVi09,PuAmKa09,TaCrEtViBeSa10})
a continuous formulation is sufficient because material property contrasts
are not drastic and thus conforming mesh doubling bricks can efficiently
handle mesh size variations \citep{KoTr02a,KoLiTrSuStSh04,LeChLiKoHuTr08,LeChKoHuTr09,LeKoHuTr09}.\newline

For a detailed introduction to the SEM as applied to regional seismic
wave propagation, please consult \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,TrKoLi08,KoVi98,KoTr99,ChKoViCaVaFe07}
and in particular \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04,KoLiTrSuStSh04}.
A detailed theoretical analysis of the dispersion
and stability properties of the SEM is available in \citet{Coh02}, \citet{DeSe07}, \citet{SeOl07}, \citet{SeOl08} and \citet{MeStTh12}.\newline

Effects due to lateral variations in compressional-wave speed, shear-wave
speed, density, a 3D crustal model, topography and bathymetry are
included. The package can accommodate full 21-parameter anisotropy
(see~\citet{ChTr07}) as well as lateral variations in attenuation
\citep{SaKoTr10}. Adjoint capabilities and finite-frequency kernel
simulations are included \citep{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,LiTr06,FiIgBuKe09,ViOp09}.\newline

The SEM was originally developed in computational fluid dynamics \citep{Pat84,MaPa89}
and has been successfully adapted to address problems in seismic wave
propagation. Early seismic wave propagation applications of the SEM,
utilizing Legendre basis functions and a perfectly diagonal mass matrix,
include \citet{CoJoTo93}, \citet{Kom97}, \citet{FaMaPaQu97}, \citet{CaGa97},
\citet{KoVi98} and \citet{KoTr99}, whereas applications involving
Chebyshev basis functions and a non-diagonal mass matrix include \citet{SePr94},
\citet{PrCaSe94} and \citet{SePrPr95}.
In the Legendre version that we use in SPECFEM the mass matrix is purposely slightly inexact but diagonal (but can be made exact if needed, see \cite{Teu15}),
while in the Chebyshev version it is exact but non diagonal.\newline

Beware that, in a spectral-element method, some spurious modes (that have some similarities with classical so-called "Hourglass modes" in finite-element techniques,
although in the SEM they are not zero-energy modes) can appear in some (but not all) cases in the spectral element in which the source is located.
Fortunately, they do not propagate away from the source element.
However, this means that if you put a receiver in the same spectral element as a source, the recorded signals may in some cases be wrong, typically exhibiting some spurious
oscillations, which are often even non causal.
If that is the case, an easy option is to slightly change the mesh in the source region in order to get rid of these Hourglass-like spurious modes,
as explained in \cite{DuLiScGa14}, in which this phenomenon is described in details, and in which practical solutions to avoid it are suggested.\newline

All SPECFEM3D software is written in Fortran2003 with full portability
in mind, and conforms strictly to the Fortran2003 standard. It uses
no obsolete or obsolescent features of Fortran. The package uses parallel
programming based upon the Message Passing Interface (MPI) \citep{GrLuSk94,Pac97}.\newline

SPECFEM3D won the Gordon Bell award for best performance at the SuperComputing~2003
conference in Phoenix, Arizona (USA) (see \citet{KoTsChTr03} and
\url{www.sc-conference.org/sc2003/nrfinalaward.html}). It was a finalist
again in 2008 for a run at 0.16 petaflops (sustained) on 149,784 processors
of the `Jaguar' Cray XT5 system at Oak Ridge National Laboratories
(USA) \citep{CaKoLaTiMiLeSnTr08}. It also won the BULL Joseph Fourier
supercomputing award in 2010.\newline

It reached the sustained one petaflop performance level for the first time in February 2013
on the Blue Waters Cray supercomputer at the National Center for Supercomputing Applications (NCSA), located at the University of Illinois at Urbana-Champaign (USA).\newline

This new release of the code includes Convolution or
Auxiliary Differential Equation Perfectly Matched absorbing Layers
(C-PML or ADE-PML) \citep{MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10,KoMa07}.
It also includes support for GPU graphics card acceleration \citep{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.\newline

The next release of the code will use the PT-SCOTCH parallel and threaded
version of SCOTCH for mesh partitioning instead of the serial version.\newline

@@

\subsection*{Announcements}
\begin{itemize}
\item {\bf FWI}: SPECFEM3D can now perform full waveform inversion (FWI), i.e. invert for models in an iterative fashion, and it can also perform source inversions in a constant model;
please refer to the two new directories, \texttt{inverse\_problem\_for\_model} and \texttt{inverse\_problem\_for\_source}, and the README files they contain. For FWI inversions for the model, also refer to the new examples provided in
directory \texttt{EXAMPLES}.\newline

\item {\bf External coupling}: SPECFEM3D can now perform coupling with an external code (DSM, AxiSEM or FK) based on a database of displacement vectors and traction vectors
on the outer edges of the mesh created once and for all (see \cite{MoChKoFu13,MoChKoWa15,WaChMoKoMo16,ToChKoBaLi14,ToKoTsHuChBaLi14}, and if you use that feature of the code please cite at least one of these papers). \newline

To use coupling with FK, just use the set of parameters that is in the \texttt{DATA/Par\_file} input file of the code: \newline
%
\begin{verbatim}
#-----------------------------------------------------------
#
# Coupling with an injection technique (DSM, AxiSEM, or FK)
#
#-----------------------------------------------------------
COUPLE_WITH_INJECTION_TECHNIQUE = .false.
INJECTION_TECHNIQUE_TYPE        = 3       # 1 = DSM, 2 = AxiSEM, 3 = FK
MESH_A_CHUNK_OF_THE_EARTH       = .false.
TRACTION_PATH                   = ./DATA/AxiSEM_tractions/3/
FKMODEL_FILE                    = FKmodel
RECIPROCITY_AND_KH_INTEGRAL     = .false. # does not work yet
\end{verbatim}
That part (coupling with FK) is actively maintained and works fine. See e.g. \href{https://doi.org/10.1002/2014GL061644}{GRL\_Ping\_Tong\_2014} for some examples.
There is also an example that is provided with the code:\newline
\texttt{specfem3d/EXAMPLES/small\_example\_coupling\_FK\_specfem}. \newline

Regarding coupling with DSM, that part is not actively maintained any more, but it is still included in the code, you may have to test it again and make minor adjustments if needed. The necessary tools are in directory \texttt{specfem3d/EXTERNAL\_PACKAGES\_coupled\_with\_SPECFEM3D/DSM\_for\_SPECFEM3D}, and there is a \texttt{README} file in \texttt{specfem3d/EXTERNAL\_PACKAGES\_coupled\_with\_SPECFEM3D} that should be more or less up-to-date (there are about four steps to follow in total, the first one being creating the database of DSM tractions and displacements on the edges of the coupling box).
See e.g. \href{https://doi.org/10.1093/gji/ggs006}{GJI\_Vadim\_2013} and \href{https://doi.org/10.1093/gji/ggv189}{GJI\_Vadim\_2015} for some examples.\newline

\item {\bf Gravity}: SPECFEM3D can now perform gravity field calculations in addition (or instead of) seismic wave propagation only.
See flag \texttt{GRAVITY\_INTEGRALS} in file \texttt{setup/constants.h}. Please also refer to \href{https://doi.org/10.1093/gji/ggx010}{GJI\_Martin\_gravimetry\_2017}.
And yes, that is the reason why Dimitri added a falling apple on the cover of the manual :-).
Note that SPECFEM3D can also model transient gravity perturbations induced by earthquake rupture, as developed and explained in \cite{HaAmBaChMoSoWh2015}.
These are two different things, and both are implemented and avaible in SPECFEM3D.
To use the second feature, please refer to\newline
\texttt{doc/how\_to/README\_gravity\_perturbation.txt}.\newline

@@

You can find all the references below in \BibTeX format in file \texttt{doc/USER\_MANUAL/bibliography.bib}.\newline

If you use SPECFEM3D Cartesian for your own research, please cite
at least one of the following articles:
\begin{description}
\item [{Numerical simulations in general}] ~\newline
 Forward and adjoint simulations are described in detail in \citet{TrKoLi08,PeKoLuMaLeCaLeMaLiBlNiBaTr11,VaCaSaKoVi99,KoMiEr09,KoErGoMi10,ChKoViCaVaFe07,MaKoDi09,KoViCh10,CaKoLaTiMiLeSnTr08,TrKoHjLiZhPeBoMcFrTrHu10,KoRiTr02,KoTr02a,KoTr02b,KoTr99}
or \citet{KoVi98}. Additional aspects of adjoint simulations are
described in \citet{TrTaLi05,LiTr06,TrKoLi08,LiTr08,TrKoHjLiZhPeBoMcFrTrHu10,PeKoLuMaLeCaLeMaLiBlNiBaTr11}.
Domain decomposition is explained in detail in \citet{MaKoBlLe08},
and excellent scaling up to 150,000 processor cores is shown for instance
in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11},
\item [{GPU computing}] ~\newline
 Computing on GPU graphics cards for acoustic or seismic wave propagation
applications is described in detail in \citet{Kom11,MiKo10,KoMiEr09,KoErGoMi10}.
\end{description}
\noindent If you use this new version, which has non blocking
MPI for much better performance for medium or large runs, please cite
at least one of these six articles, in which results of non blocking
MPI runs are presented: \citet{PeKoLuMaLeCaLeMaLiBlNiBaTr11,KoErGoMi10,KoViCh10,Kom11,CaKoLaTiMiLeSnTr08,MaKoBlLe08}.\newline

@@

\noindent If you work on geophysical applications, you may be interested
in citing some of these application articles as well, among others:
\begin{description}
\item [{Southern California simulations}] ~\newline
 \citet{KoLiTrSuStSh04,KrJiKoTr06a,KrJiKoTr06b}.

@@

\item [{Anisotropy}] ~\newline
 \citet{ChTr07,JiTsKoTr05,ChFaKo04,FaChKo04,RiRiKoTrHe02,TrKo00}.
\item [{Attenuation}] ~\newline
 \citet{SaKoTr10,KoTr02a,KoTr99}.
\item [{Topography}] ~\newline
 \citet{LeKoHuTr09,LeChKoHuTr09,LeChLiKoHuTr08,GoAmTaCaSmSaMaKo09,WiKoScTr04}.
\end{description}
The corresponding Bib\TeX{} entries may be found in file \texttt{doc/USER\_MANUAL/bibliography.bib}.
